---
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

Load the required packages and set a seed for the random number generator.
```{r}
library(tidyverse)

library(rstan)
# set cores to use to the total number of cores (minimally 4)
options(mc.cores = max(parallel::detectCores(), 4))
# save a compiled version of the Stan model file
rstan_options(auto_write = TRUE)

library(brms)

# install faintr with devtools::install_github('michael-franke/bayes_mixed_regression_tutorial/faintr', build_vignettes = TRUE)
library(faintr)

set.seed(123)

```

Read the data:
```{r}
raw_data = read_csv("results.csv")
glimpse(raw_data)
```  

Rename variables
```{r}
data = rename(raw_data,
  submissionID = `submission_id`,
  trialName = `trial_name`
)

glimpse(data)
```

Select only the variables that are important for the analysis:
```{r}
data = select(data, submissionID, trialName, 
         response, pictureNumber, artist, date, 
         age, gender, timeSpent, color, color2)

glimpse(data)

```

Exclude participants from the data that we don't want to have. 
Also find out why we had to exclude them.
```{r}
data_ex <- data

fehlerCV1 = 0
fehlerCV2 = 0
fehlerV = 0
fehlerExp = 0
fehlerT = 0
fehlerXP = 0


for(i in 1:(nrow(data) / 245)) {
  
  # colorVisionTest1
  if(data[((i - 1)*245 + 1), ]$response != "74") {
    data_ex = filter(data_ex, submissionID != data[((i - 1)*245 + 1), ]$submissionID)
    
    fehlerCV1 = fehlerCV1 + 1
    
  # colorVisionTest2
  } else if(data[((i - 1)*245 + 2), ]$response != "7") {
    data_ex = filter(data_ex, submissionID != data[((i - 1)*245 + 2), ]$submissionID)
    
     fehlerCV2 = fehlerCV2 + 1
     
  # visionTest
  } else if(data[((i - 1)*245 + 3), ]$response != "FELOPZD") {
    data_ex = filter(data_ex, submissionID != data[((i - 1)*245 + 3), ]$submissionID)
    
     fehlerV = fehlerV + 1
     
  # expertise 
  } else if(as.integer(data[((i - 1)*245 + 244), ]$response) > 5) {
     data_ex = filter(data_ex, submissionID != data[((i - 1)*245 + 244), ]$submissionID)
   
      fehlerExp = fehlerExp + 1
      
  # time spent 
  } else if(data[((i - 1)*245 + 1), ]$timeSpent < 5) {
     data_ex = filter(data_ex, submissionID != data[((i - 1)*245 + 1), ]$submissionID)
   
      fehlerT = fehlerT + 1
      
  # exclude participants that took part in the course Experimental Psychology Lab
  } else if(data[((i - 1)*245 + 245), ]$response == "Yes") {
    data_ex = filter(data_ex, submissionID != data[((i - 1)*245 + 245), ]$submissionID)
    
     fehlerXP = fehlerXP + 1
     
  }
  
}

glimpse(data_ex)

print(c("Number of excluded participants because of colorVisionTest1: ", fehlerCV1))
print(c("Number of excluded participants because of colorVisionTest2: ", fehlerCV2))
print(c("Number of excluded participants because of VisionTest: ", fehlerV))
print(c("Number of excluded participants because of their expertise level in cubism: ", fehlerExp))
print(c("Number of excluded participants because of TimeSpent: ", fehlerT))
print(c("Number of excluded participants because of participation in XPLab: ", fehlerXP))

```

Calculate age mean, age range and gender distribution, number of participants:
```{r}
data_ex %>% summarise(meanAge = mean(na.exclude(age)))

gender_distribution = count(data_ex, gender) 
gender_distribution$n =  gender_distribution$n / 245

print(c("Age range: ", range(na.exclude(data_ex$age))))

participants = nrow(data) / 245
participants_valid = nrow(data_ex) / 245
    
print(gender_distribution)
print(c("Total number of participants: ", participants))
print(c("Number of valid participants: ", participants_valid))

```

transform the data into right form: 
```{r}
d = filter(data_ex, trialName %in% c("ratingScaleLike", 
                                 "ratingScaleDetect")) %>%
  select(submissionID, trialName, pictureNumber, artist, date, response, color, color2) %>% 
  spread(key = trialName, value = response) 

d$ratingScaleLike <- as.numeric(d$ratingScaleLike)
d$ratingScaleDetect <- as.numeric(d$ratingScaleDetect)

d = rename(d, detectability = ratingScaleDetect, liking = ratingScaleLike)

glimpse(d)
```

Check that everything is there and of the correct type:
```{r}
d_tibble <- as_tibble(d)

print(d_tibble)

```

inspect the data with x = detectability and y = liking for the different participants
This part is just for our interest.
```{r}
ggplot(d, aes(x = detectability, y = liking, group = submissionID)) + geom_jitter() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking")

```

inspect the data x = detectability and y = liking
```{r}
ggplot(d, aes(x = detectability, y = liking)) + geom_jitter() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking")
ggplot(d, aes(x = detectability, y = liking)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking")
```

Boxplots that shows liking and detectability 
```{r}
ggplot(d, aes(x = detectability, y = liking, group = detectability)) + geom_jitter() + geom_boxplot() + labs(x = "Detectability", y = "Liking")
ggplot(d, aes(x = detectability, y = liking, group = detectability)) + geom_boxplot() + labs(x = "Detectability", y = "Liking")
```

calculate means for detectability:
```{r}
d %>% group_by(liking) %>% summarise(meanDetect = mean(detectability))
```

calculate means for liking:
```{r}
d %>% group_by(detectability) %>% summarise(meanLike = mean(liking))
```

calculate means for different artists:
```{r}
d %>% group_by(artist) %>% summarise(meanLike = mean(liking), meanDetect = mean(detectability))

```

Boxplots and graphs that show liking and detectability (for the different artists)
```{r}
dataB <- d %>% filter(artist == "B")
dataP <- d %>% filter(artist == "P")
dataG <- d %>% filter(artist == "G")

ggplot(dataB, aes(x = detectability, y = liking, group = detectability)) + geom_boxplot() + labs(x = "Detectability", y = "Liking", title = "Braque")
ggplot(dataP, aes(x = detectability, y = liking, group = detectability)) + geom_boxplot() + labs(x = "Detectability", y = "Liking", title = "Picasso")
ggplot(dataG, aes(x = detectability, y = liking, group = detectability)) + geom_boxplot() + labs(x = "Detectability", y = "Liking", title = "Gris")

ggplot(dataB, aes(x = detectability, y = liking)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking", title = "Braque")
ggplot(dataP, aes(x = detectability, y = liking)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking", title = "Picasso")
ggplot(dataG, aes(x = detectability, y = liking)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking", title = "Gris")

```

Graphs / boxplots that show liking and detectability (for the different artists in one graph / boxplot)
```{r}
ggplot(d, aes(x = detectability, y = liking, fill = artist)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking", title = "Different artists", fill = "Artist")

ggplot(d, aes(x = as.factor(detectability), y = liking, fill = artist)) + geom_boxplot()  + labs(x = "Detectability", y = "Liking", title = "Different artists", fill = "Artist")

```

Graphs / boxplots that show liking and detectability (for the color vs. monochrome in one graph / boxplot)
```{r}
ggplot(d, aes(x = detectability, y = liking, fill = color)) + geom_point() + geom_smooth(method = "lm" ) + labs(x = "Detectability", y = "Liking", title = "Different colors", fill = "Color")

ggplot(d, aes(x = as.factor(detectability), y = liking, fill = color)) + geom_boxplot() + labs(x = "Detectability", y = "Liking", title = "Different colors", fill = "Color")

```

Graphs / boxplots that show liking and detectability (for the color vs. monochrome vs. sepia in one graph / boxplot)
```{r}
ggplot(d, aes(x = detectability, y = liking, fill = color2)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking", title = "Different colors", fill = "Color")

ggplot(d, aes(x = as.factor(detectability), y = liking, fill = color2)) + geom_boxplot() + labs(x = "Detectability", y = "Liking", title = "Different colors", fill = "Color")

```

calculate means for different colors:
```{r}
d %>% group_by(color2) %>% summarise(meanLike = mean(liking))

```

Boxplots and graphs that show liking and detectability (for the different time span)
```{r}
data1 <- d %>% filter(date <= 1911)
data2 <- d %>% filter(date >= 1912, date <= 1915)
data3 <- d %>% filter(date >= 1916)

ggplot(data1, aes(x = detectability, y = liking, group = detectability)) + geom_boxplot() + labs(x = "Detectability", y = "Liking", title = "Paintings from 1908-1911")
ggplot(data2, aes(x = detectability, y = liking, group = detectability)) + geom_boxplot() + labs(x = "Detectability", y = "Liking", title = "Paintings from 1912-1915")
ggplot(data3, aes(x = detectability, y = liking, group = detectability)) + geom_boxplot() + labs(x = "Detectability", y = "Liking", title = "Paintings from 1916-1919")

ggplot(data1, aes(x = detectability, y = liking)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking", title = "Paintings from 1908-1911")
ggplot(data2, aes(x = detectability, y = liking)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking", title = "Paintings from 1912-1915")
ggplot(data3, aes(x = detectability, y = liking)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Detectability", y = "Liking", title = "Paintings from 1916-1919")

```

prepare the data for the analysis:
```{r}
d = mutate(d, 
           liking = factor(liking, ordered = T),
           detectability = factor(detectability, ordered = T),
           color = factor(color),
           color2 = factor(color2),
           artist = factor(artist)
           )
glimpse(d)
```

ANALYSIS

Pearson correlation of liking and detectability: 
```{r}
cor(as.numeric(d$liking), as.numeric(d$detectability))
```

M1. simple analysis: Treat the detectability scale as a factor.
Calculate the difference of the sample points: compare the sample points for each category of detectability (2 to 7) to the sample points from detectability = 1. 
```{r}
m1 = brm(formula = liking ~ as.numeric(detectability),
         data = d,
         family = cumulative("logit"))

print(m1)

marginal_effects(m1, categorical = T)
marginal_effects(m1, categorical = F)
```

M2: Model that treats detectability as monotonic and includes random effects
```{r}

m2 = brm(formula = liking ~ mo(detectability) + (1|submissionID),
         data = d,
         family = "cumulative")

print(m2)

marginal_effects(m2, categorical = T)
marginal_effects(m2, categorical = F)

```

M3: Model that treats detectability as monotonic
```{r}

m3 = brm(formula = liking ~ mo(detectability),
         data = d,
         family = "cumulative")

print(m3)

marginal_effects(m3, categorical = T)
marginal_effects(m3, categorical = F)

```

M4: Model that additionlly includes artist and color as an independent variable and includes random effects
```{r}

m4 = brm(formula = liking ~ mo(detectability) + artist + color2 + (1|submissionID), 
         data = d,
         family = "cumulative")

print(m4)

marginal_effects(m4, categorical = T)
marginal_effects(m4, categorical = F)
```

M5: Like M4, but without random effects
```{r}

m5 = brm(formula = liking ~ mo(detectability) + artist + color2, 
          data = d,
          family = "cumulative")

print(m5)

marginal_effects(m5, categorical = T)
marginal_effects(m5, categorical = F)
```

Save current workspace to be able to compare which model fits best afterwards
```{r}
save.image("models.RData")
```

Analyse which model fits our data best
```{r}
load("models.RData")

m1 <- add_criterion(m1, "waic")
m2 <- add_criterion(m2, "waic")
m3 <- add_criterion(m3, "waic")
m4 <- add_criterion(m4, "waic")
m5 <- add_criterion(m5, "waic")

loo_compare(m1, m2, m3, m4, m5, criterion = "waic")

```
